---
title: "hw-04"
author: "Guochenchen"
date: "October 7th, 2018"
output: github_document
always_allow_html: yes
---


#Homework 04: Tidy data and joins

## Data Reshaping Prompts(And relationship to aggregation)

```{r}
# Firstly load all libraries
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(wbstats))

```
### 1. Make a tibble with one row per year and columns for life expectancy for two or more countries
```{r}
data1 <- gapminder %>% 
  filter(country %in% c('China', 'Japan', 'Canada')) %>% 
  select(year, country, lifeExp)

knitr::kable(head(data1, 20)) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
  

# To further explore this table
data2 <- spread(data1, key = "country", value = "lifeExp") %>% 
  rename(lifeexp_Canada = Canada, lifeexp_China = China, lifeexp_Japan = Japan)

knitr::kable(data2) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

#To scatter plot this data further indicate the comparison between lifeexp of these three countries
ggplot(data1, aes(year, lifeExp)) +
  geom_point(aes(color = country))+
  scale_x_continuous(limits = c(1952, 2007), breaks = seq(1952, 2007, 5))+
  xlab("Year")+ ylab("Life expectancy") + ggtitle("Life expectancy of Canada, China and Japan")
  
```

### 2. Compute life expectancy for all possible combinations of content and year. Reshape that to have one row per year and one variable for each continent.
```{r}
# Here to compute the maximum lifeExpectancy for all continents each year
data3 <- gapminder %>% 
  group_by(continent, year) %>% 
  summarise(MaxLifeExp = max(lifeExp))

data4 <- spread(data3, key = "continent", value = "MaxLifeExp")

knitr::kable(head(data4, 20)) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# alter the x, y of table 
data5 <- spread(data3, key = "year", value = "MaxLifeExp")

knitr::kable(head(data5, 20)) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# And the plot of data 4
ggplot(data = data3, aes(x=year, y=MaxLifeExp, group = continent, colour = continent))+
  geom_line(size = 1)+
  scale_x_continuous(limits = c(1952, 2007), breaks = seq(1952, 2007, 5))+
  xlab("Year")+ ylab("Max Life Expectancy") + ggtitle("Maximum Life expectancy tendency of each continent from 1952 to 2007")

# another way to show this, use bar chart to separately indicate the maximum lifeexpectancy of each continent
ggplot(data = data3, aes(x=continent, y=MaxLifeExp, group=continent))+
  facet_wrap( ~year)+
  geom_col(aes(fill=MaxLifeExp), size = 5)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  ggtitle("Maximum Life expectancy of each continent from 1952 to 2007")
  
```
### 3. Reshape the table to have one row per year or per year * continent combination
```{r}
data6 <- gapminder %>% 
  group_by(year, continent) %>% 
  filter(min_rank(lifeExp)<2 | min_rank(desc(lifeExp))<2) %>% 
  select(year, continent, country, lifeExp)
knitr::kable(head(data6, 20)) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# for per year*continent combination
data7 <- data6 %>% 
  group_by(year, continent) %>% 
  arrange(lifeExp) %>% 
  mutate(MinlifeExp_country = first(country), 
         MinlifeExp_value = min(lifeExp),
         MaxlifeExp_country = last(country),
         MaxlifeExp_value = max(lifeExp)) %>% 
  arrange(year) %>% 
  unite("Year_Cont", year, continent) %>% 
  select(Year_Cont, MinlifeExp_country, MinlifeExp_value, MaxlifeExp_country, MaxlifeExp_value)
# Form to table
knitr::kable(data7) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
```
### 4. Data manipulation sampler
```{r}
# filter out gdp for china, japan and canada for each year
data8 <- gapminder %>% 
  filter(country %in% c("China", "Japan", "Canada")) %>% 
  select(year, country, gdpPercap)
# only show 15 of whole data to keep markdown file compact
knitr::kable(head(data8, 15)) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# calculate the total gdp of three countries for each year
data9 <- data8 %>% 
  group_by(year) %>% 
  summarize(Total_country=n(),
            Total_gdp=sum(gdpPercap))
knitr::kable(data9) %>% 
  kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# the gapminder dataset is not a perfect example to show dcast and with function here.
C <- with(data9, table(year, Total_country))
kable(C)
```
## Join Prompts(join, merge, look up)
### Load another data package: country_Capitals
```{r}
# The second data frame used complementary to gapminder is country_capitals.

coun_cap <- read.csv("https://github.com/icyrockcom/country-capitals/raw/master/data/country-list.csv")
ls(coun_cap)
nrow(coun_cap)
nrow(gapminder)
# Show the country contained in country_capitals but not in gapminder
setdiff(coun_cap$country, gapminder$country)
# show countries contained in gapminder but not in country_capitals
setdiff(gapminder$country, coun_cap$country)

```
### Activity 1

```{r}
head(coun_cap)

# use left join to join these two data
coun_cap$type <- NULL
# To delete the type column of package country capital which will not be used in this join function.
# And firstly left join gapminder and capitals by country column
join1 <- gapminder %>% 
  filter(year == "2007") %>% 
  left_join(coun_cap, by = "country")
knitr::kable(join1) %>% 
   kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

#Then right join, the difference are right join will contain all countries in countrycapitals, hence lots ofcountries has no values of continent, year, lifeExp and etc.
join2 <- gapminder %>% 
  filter(year == "2007") %>% 
  right_join(coun_cap, by = "country")
knitr::kable(join2) %>% 
   kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

# Then use inner join, this time only intersect datas will be shown
join3 <- gapminder %>% 
  filter(year == "2007") %>% 
  inner_join(coun_cap, by = "country")
knitr::kable(join3) %>% 
   kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
# And semi join, here only countries of gapminder that in country capitals will be displayed. this means some counties that gapminder has but not in country capitals will be ignored
join4 <- gapminder %>% 
  filter(year == "2007") %>% 
  semi_join(coun_cap, by = "country")
knitr::kable(join4) %>% 
   kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)

#Anti-join: This will display those countries in gapminder but not in country capitals
join5 <- gapminder %>% 
  filter(year == "2007") %>% 
  anti_join(coun_cap, by = "country")
knitr::kable(join5) %>% 
   kable_styling(bootstrap_options = "bordered", latex_options = "basic", full_width = F)
